### 2020-12-02 Non-converging-case-study.Rmd

## Case Study

- We use real data: the `boys` dataset from the `mice` package

- We are interested in predicting age from the other variables, in particular in the regression coefficient of `hgt`

- We compare non-convergence identified using visual inspection versus rhat in the chain variances, scientific estimate and lambda.

- The figures show results of a `mice` run with 30 iterations but otherwise default settings.

```{r echo=FALSE}
# autocorrelation plots (not used)
# casestudy_plot(v="ac.max.qhat") #+ list(scale_y_continuous(limits = c(0.65,1)))
# casestudy_plot(v="ac.max.lambda")
```

```{r echo=FALSE}
p <- plot(mids)
p[1:2,1] # %>% plotly::ggplotly()
```

From the traceplot of the chain means it seems that mixing improves up-to 20 iterations, and there is some trending up-to 20 iterations as well. The chain variances are more or less stationary after 5 iterations, and mixing requires about 20 iterations.

```{r echo=FALSE}
# p <- casestudy_plot(v="est", name = "Regression estimate")
# p %>% plotly::ggplotly()
```

```{r echo=FALSE}
p <- casestudy_plot(v="r.hat.max.mu.hgt", name = "Rhat chain means") + 
  list(
    geom_hline(yintercept = 1.2, color = "grey")#, 
    #scale_y_continuous(limits = c(1,5))
    )
p # %>% plotly::ggplotly()
```

This figure shows that 36 iterations are required before the $\widehat{R}$-values of the chain means drop below the threshold for non-convergence.

```{r echo=FALSE}
p <-
  casestudy_plot(v = "r.hat.max.qhat", name = "Rhat regression estimate") +
  list(geom_hline(yintercept = 1.2, color = "grey")#,
       #scale_y_continuous(limits = c(1, 2.5))
       )
p # %>% plotly::ggplotly()
```

The scientific estimate reaches the threshold much sooner, when $it=15$.

```{r echo=FALSE}
p <- casestudy_plot(v = "r.hat.max.lambda", name = "Rhat lambda") +
  list(geom_hline(yintercept = 1.2, color = "grey")#,
       #scale_y_continuous(limits = c(1, 2.5))
       )
p # %>% plotly::ggplotly()
```

According to the $\widehat{R}$-values of the novel parameter, at least 25 iterations are required.

[Is this a nice example? Why is convergence so slow? Don't the defaults work? Is this the bmi+wgt+wgt problem? Maybe find an easier problem?]